---
title: "HW1: Part 1"
output: html_document
---
# Question 1: New York Times Data

```{r}
nyt1 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt1.csv")
nyt1 <- na.omit(nyt1)
str(nyt1)
summary(nyt1$Age)
dim(nyt1)
```

##Part 1: Create a new variable, age_group, that categorizes users as “<18”, ”18-24”, ”25-34”, ”35-44”, ”45-54”, “55-64” and “65+”. Initially I ran if, else f and else loop inside a for loop for the number of rows in the dataframe but then later saw the cut function and used that 

```{r}
breaks <- c(0, 18, 25, 35, 45, 55, 65, 110) #Summary commands tells us that 108 is the highest age
labels <- c('<18', '18-24', '25-34', '35-44', '45-54', '55-64', '64+')
nyt1$Age_Group <- cut(nyt1$Age, breaks=breaks, labels=labels, right = FALSE, ordered_result = TRUE)
summary(nyt1$Age_Group) # to ensure there are no null values
```

## Part 2.1: For a single day plot the distributions of number impressions and click-through-rate (CTR=# clicks/# impressions), for these 6 age categories.

```{r}
nyt1$ctr <- nyt1$Clicks/nyt1$Impressions
summary(nyt1$ctr)
nyt1$ctr[is.na(nyt1$ctr)] <- 0 # Replacing NA values with 0 
```

```{r}
library(ggplot2)
impressions_age_group=ggplot(nyt1, aes(x=Impressions, colour=Age_Group)) + geom_density() + ggtitle("Number of Impressions by Age Group")
plot(impressions_age_group)
ctr_age_group=ggplot(nyt1, aes(x=ctr, colour=Age_Group)) + geom_density() + ggtitle("Number of Click-Through-Rate by Age Group")
plot(ctr_age_group)
```

## Part 2.2: Define a new variable to segment or categorize users based on their click behavior 
```{r}
table1 <- prop.table(table(nyt1$Clicks))*100 #gives percentage distribution of clicks values
table1
### Creating a variable called have_clicked where if Clicks >=1 the value will be 1 else 0
nyt1$have_clicked <- ifelse(nyt1$Clicks>=1, 1, 0) 
table_have_clicked <- prop.table(table(as.factor(nyt1$have_clicked)))
table_have_clicked
```

## Part 2.3: Explore the data and make visual and quantitative comparisons across user segments/ demographics (<18 year old male vs < 18 year old females or logged-in vs not, for example).

```{r}
table1 <- table(nyt1$have_clicked, nyt1$Gender)
barplot(table1, beside=T, legend.text=c("Not Clicked", "Clicked") , main="Clicked Behavior Categorized by Gender", xlab="Gender", ylab="Have Clicked", col=c(2,4))
```

###Relative frequency of Clicks by age Group

```{r}
library("plyr")
Clicked <- prop.table(table(nyt1$Age_Group, nyt1$have_clicked==1))  
Clicked <- data.frame(Clicked)
Clicked <- subset(Clicked, Var2==TRUE) 
Clicked$Var2 <- NULL
```
Clicked<- rename(Clicked,c("Var1" = "AgeGroup"))
Clicked
```{r}
library(data.table)
s <- sum(Clicked$Freq)
head(s)
Clicked$RelFrq <- Clicked$Freq/s    #calculate relative frequency
labl <- data.table(Clicked)[, per := sprintf("%.1f%%", RelFrq*100), by = Var1]
head(labl) 
```

```{r}
have_clicked_age_group=ggplot(Clicked, aes(x=Var1, y=RelFrq, fill=Var1)) + geom_bar(stat="identity") + labs(title="At least 1 Clicks by Age Group - Day 1", x= "Age Groups", y= "Relative Frequency of Clicks")+ guides(fill=FALSE)+ geom_text(data = labl, aes(x = Var1, y = RelFrq, label = per), vjust= -0.5)
plot(have_clicked_age_group)

```

###Now let's check average Impressions by gender and age group

```{r}
means <- with(nyt1, aggregate(x=list(Y=Impressions),by=list(A=Age_Group, B=Gender),mean)) 
with(means, interaction.plot(x.factor=A, trace.factor=B, response=Y, type='b', main = "Impressions by gender and age group, day 10", xlab = "Age groups", ylab= "Average Impressions"))

#This shows an interesting result - generally, the average number of impressions decreases as you move up in age group. Women, ‘0’, seem to have a higher average of impressions with every age group.
```

#Now let’s check out if there are differences with gender and age group with clicks…

```{r}
means1 <- with(nyt1, aggregate(x=list(Y=Clicks), by=list(A=Age_Group, B=Gender),mean))
with(means1, interaction.plot(x.factor=A, trace.factor=B, response=Y, type='b',main = "Clicks by gender and age group, day 10", xlab = "Age groups", ylab= "Average Clicks"))
```

##Part 3: Create & visualize metrics/measurements/statistics that summarize the data. 

```{r}
#We will calculate average Impressions of page across Age Group
nyt1$mean_Imp<-nyt1$Impressions
mean_Impression_AgeGroup <- ddply(nyt1, "Age_Group", summarise, mean_Imp.mean=mean(mean_Imp))
mean_Impression_AgeGroup

#Average Click Through Rate across Age Groups
mean_ctr <- ddply(nyt1, "Age_Group", summarise, ctr.mean=mean(ctr))
mean_ctr

#Average Impressions across Gender
nyt1$AvgeImp<-nyt1$Impressions
mean_Imp_Gender <- ddply(nyt1, "Gender", summarise, ImpressionByGender.mean=mean(AvgeImp))
mean_Imp_Gender
```

##Now we will extend our analysis across one week.

```{r}
nyt1_new <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt1.csv")
nyt2 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt2.csv")
nyt3 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt3.csv")
nyt4 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt4.csv")
nyt5 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt5.csv")
nyt6 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt6.csv")
nyt7 <- read.csv("/Users/mayankajha/Documents/Spring-2018/INST_737/nyt7.csv")

nyt1_new$day <- "Day 1"
nyt2$day <- "Day 2"
nyt3$day <- "Day 3"
nyt4$day <- "Day 4"
nyt5$day <- "Day 5"
nyt6$day <- "Day 6"
nyt7$day <- "Day 7"

nyt <- rbind(nyt1_new, nyt2, nyt3, nyt4, nyt5, nyt6, nyt7)
#View(nyt)

levels(nyt$day) <- list('day1' = 'day1',
                        'day2' = 'day2',
                        'day3' = 'day3',
                        'day4' = 'day4',
                        'day5' = 'day5',
                        'day6' = 'day6',
                        'day7' = 'day7')

#encoding Female as 0 and Male as 1
nyt$Gender <- ifelse(nyt$Gender == 0, 'Female','Male') 

#calculating click through rate
nyt$CTR <- ifelse(nyt$Impressions != 0, nyt$Clicks/nyt$Impressions,NA) 
```


###Comparing mean number of clicks/person for both genders across 7 days
```{r}
clicks.plot2 <- aggregate(nyt$Clicks, by = list(nyt$day, nyt$Gender), FUN = mean, na.rm = TRUE)

#From table  we can observe below points: 
#1. Average number of clicks/ person does not vary throug day1 to day7 for either male or female. 
#2. Female has higher average number of clicks/ person when compared to male
clicks.plot2
```

###Comparing mean CTR for both genders across 7 days
```{r}
ctr.plot3 <-aggregate(nyt$CTR,
                      by = list(nyt$day, nyt$Gender),
                      FUN = mean, na.rm = TRUE)

#From table  we can observe below points:
#1. Average CTR/ person does not vary throug day1 to day7 for either male or female.  
#2. Female has higher average number of CTR/ person when compared to male 
ctr.plot3
```

###Plotting distribution of Clicks and CTR for Males and Females for Day1 to Day7
```{r}
#From plot we can observe that clicks remain amost same for all days except for day 5 and day 6. On day 5 it is lower than normal and on day 6 it s above normal level. The behavior is same for both male and female

agg = ggplot(nyt, aes(day, Clicks))+ geom_bar(stat = "identity") + facet_grid(.~ Gender)
plot(agg)
agg1 = ggplot(nyt, aes(day, CTR))+ geom_bar(stat = "identity") + facet_grid(.~ Gender)
plot(agg1)
```

#Part 4: Describe and interpret any patterns you find.

1. From plot we observed that clicks remain almost same for all days except for day 5 and day 6
   + On day 5 it is lower than normal 
   + On day 6 it s above normal level 
2. It is also clear that people from the age group less than 18 years are most likely (and most frequently) to click than any other age group.
3. It appears that females are most likely to click than male. 
4. We see no anomalies in the given data. 
   + Average number of Clicks and CTR/ person does not vary throug day1 to day7 for either male or female. 
   + However, female has higher average number of Clicks and CTR/ person when compared to male. 
